### 2014-05-07 ApTranscriptome_TR.Rmd
Updated script to remove contaminated transcripts after expression quantification

This resulted in removing 5,675 contigs as contaminants, leaving 99,861 "clean" contigs. We spot-checked the contaminants by BLAST and confirmed that they matched bacteria, human or viral sources by greater than 95%.

Running Trinity and subsequent programs is time and memory-intensive, so the final filtered clean assembly can be downloaded from [...]

~~~
# download filtered Trinity assembly, uncompress and move
wget http://johnstantongeddes.org/assets/files/Aphaenogaster_transcriptome.tar
tar -xvf Aphaenogaster_transcriptome.tar
# check md5sum
md5sum Trinity_cap3_uclust_clean.fa
# 0160a82ec62c6f371c84e4963fed09d1
mkdir -p results/
mkdir -p results/trinity-full/
mv Trinity_cap3_uclust.fasta results/trinity-full/.
mv Trinity_cap3_uclust_clean.fasta results/trinity-full/.
~~~

@@

@@

**Preliminary [examination](https://minilims1.uvm.edu/BCProject-26-Cahan/methods.html#clustering-of-samples) of the data indicated that the A22_7 and Ar_7 samples may have been switched, so I remove these values from the analysis.**

```{r RxN, echo=TRUE, eval=TRUE, cache=TRUE}
A22.TPM[,colony:="A22"]
Ar.TPM[,colony:="Ar"]
TPM.dt <- rbind(A22.TPM, Ar.TPM)
TPM.dt$colony <- as.factor(TPM.dt$colony)
str(TPM.dt)

setkey(TPM.dt, val)
TPM.dt.sub <- TPM.dt[val != 7] 
unique(TPM.dt.sub$val)
```

@@

# extract significant transcripts
setkey(TPM.dt.sub, Transcript)
sig.TPM.dt.sub <- TPM.dt.sub[signif.transcripts$Transcript]
```

>>>>

This resulted in removing 5,675 contigs as contaminants, leaving 99,861 "clean" contigs. We spot-checked the contaminants by BLAST and confirmed that they matched bacteria, human or viral sources by greater than 95%. For expression quantification, we use the full assembly to ensure that "contaminant" reads are assigned to the contaminants. After quantification, these transcripts will then be removed from further analyses.

Running Trinity and subsequent programs is time and memory-intensive so the final assembly can be downloaded and used for all further analyses. In addition, the compressed archive contains the "clean" and "contaminant" sequences after filtering with DeconSeq.

~~~
# download filtered Trinity assembly, uncompress and move
wget http://johnstantongeddes.org/assets/files/Aphaenogaster_transcriptome.tar
# check md5sum
md5sum Aphaenogaster_transcriptome.tar
# fd6dbb0b3e88e1c797c9e74611b245b2

# uncompress and move
tar -xvf Aphaenogaster_transcriptome.tar
mkdir -p results/
mkdir -p results/trinity-full/
mv Trinity_cap3_uclust.fasta results/trinity-full/.
mv Trinity_cap3_uclust_clean.fasta results/trinity-full/.
~~~

@@

**Preliminary [examination](https://minilims1.uvm.edu/BCProject-26-Cahan/methods.html#clustering-of-samples) of the data indicated that the A22_7 and Ar_7 samples may have been switched, so I remove these values from the combined expression data set for the two species.**

```{r combined_TPM, echo=TRUE, eval=TRUE, cache=TRUE}
A22.TPM[,colony:="A22"]
Ar.TPM[,colony:="Ar"]
TPM.dt <- rbind(A22.TPM, Ar.TPM)
TPM.dt$colony <- as.factor(TPM.dt$colony)
str(TPM.dt)

setkey(TPM.dt, val)
TPM.dt.sub <- TPM.dt[val != 7] 
unique(TPM.dt.sub$val)
length(unique(TPM.dt.sub$Transcript))
```

### Remove *contaminant* transcripts

```{r remove_contaminants}
# extract transcript names from fasta file rather than loading whole file
system("grep '^>' results/trinity-full/Trinity_cap3_uclust_cont.fa | cut -f 1 -d ' ' | sed -r 's/^.{1}//' > results/trinity-full/cont.list")

# read file
cont.list <- read.table("results/trinity-full/cont.list")
str(cont.list)

# remove from TPM.dt.sub
setkey(TPM.dt.sub, Transcript)
TPM.dt.sub <- TPM.dt.sub[!cont.list]
str(TPM.dt.sub)
length(unique(TPM.dt.sub$Transcript))
```

@@

@@

# extract significant transcripts
sig.TPM.dt.sub <- TPM.dt.sub[signif.transcripts$Transcript]
```
